# Object Documentation (`man/`)

Documentation in R is saved as `.Rd` files. These live separately to our R functions which live in `.R` files. However we can generate our `.Rd` files using an R package called `roxygen2` which parses the `.R` files.

We include a `roxygen` header above the function definition. Each line of the `roxygen` header starts with the symbols `#'`. This is known as a `roxygen` comment. Following this we use special tags to indicate a particular component of the help file. However the first three elements have special meaning. The first three paragraphs of the header are treated as:

* The title of the help page (short, one sentence)
* The description for the help page (brief description of the function)
* The details section which can provide much more information about the function, what it implements etc. 

Following these lines we must use the tags to distinguish elements. These tags all start with an `@` symbol. Some tags and their uses are:

| Tag | Purpose |
| --- | ------- |
| @param | Identify each of the function arguments and the corresponding help text |
| @return | Detail the output of the function |
| @author | Who wrote this function? |
| @seealso | Other functions that the user should also look at the help documentation for |
| @examples	| Code examples of running the function |
| @import/@importFrom | Indicate a package or function within a package to be imported |
| @export | Indicate that this function should be exported (i.e. made visible to the end user) |

The following is an example of how this might look for an example function.

```{r roxygen_ex}
#' Sample from a dataset	 
#'
#' This function has been designed to sample from the rows of a two 
#' dimensional data set returning all columns of the sampled rows.
#'
#' @param data The matrix or data.frame from which rows are to be sampled.
#' @param size The number of samples to take.
#' @param replace Should values be replaced? By default takes the value TRUE.
#' @param ... Any other parameters to be passed to the sample function.
#'
#' @return Returns a dataset of the same type as the input data with \code{size} 
#' rows.
#'
#' @author Nathan Eastwood <neastwood@@mango-solutions.com>
#' 
#' @examples
#' sampleFromData(airquality, 100)
#'
#' @export
sampleFromData <- function(data, size, replace = TRUE, ...) {
...
}
```

To generate the `.Rd` files, we run the `devtools::document()` function (Ctrl/Cmd + Shift + D).

## Using `roxygen2` to Document the Package

We can also generate the package documentation using `roxygen2`. For package documentation we need to include the tags `@docType` and `@name`. The header itself needs to be followed by `NULL` as all headers need to be followed by some form of R code. 

Whilst this header can go in any file, good practice is to create a separate R script which takes the package name that contains this header. In the `sparkts` package therefore, we have an R script named `sparkts.R` in the R directory containing the following:

```{r package_desc}
#' A package for calling Scala methods from the spark-ts package
"_PACKAGE"
```

We could of course expand this by adding a description etc.