# Testing Your Code

You can use `devtools::load_all()` to simulate building and loading your package. This is useful for when you have written some new code or a new feature and you want to interactively run it.

## `testthat`

There are a number of ways of writing unit tests in R but the simplest is to use the package `testthat`.

### Setting up the Test Framework

We can set up a test framework in a package using the function `use_testthat`.

```{r use_testthat}
use_testthat()
```

This will create a test directory that contains:

* A script `testthat.R` that contains all the code required to run the package tests
* A directory `testthat/` where we will save all of our test scripts

The `use_testthat` function will also update the `DESCRIPTION` file accordingly to include `testthat` as a suggested package. 

### Writing Tests

With the package `testthat` we create individual tests using functions named with the pattern `expect_*`. Some of the most common expect statements are given in the following table:

| Expectation | Description |
| ----------- | ----------- |
| expect_identical | Checks for exact quality |
| expect_equal | Checks for equality with numerical tolerance |
| expect_equivalent | More relaxed version of equals |
| expect_is | Checks than an object inherits a specified class |
| expect_matches | Compares a character vector to a regular expression |
| expect_true | Checks that the logical TRUE is returned |
| expect_false | Checks that the logical FALSE is returned |
| expect_error | Checks that an expression throws an error |
| expect_warning | Checks that an expression gives a warning |
| expect_message | Checks that an expression issues a message |

We typically want to group tests which belong to a single function together. We do this in the `test_that` function.

```{r test_that, eval = TRUE}
context("A and B have the same column names and same second column")

test_that("Test that A is the same as B", {
  A <- data.frame(x = rnorm(10), y = 1:10)
  B <- data.frame(x = rnorm(10), y = 1:10)
  
  expect_equal(colnames(A), colnames(B))
  expect_identical(A$y, B$y)
})
```

The `context` function here indicates that all of the `test_that` groups that follow can be collected together. To structure this as tests for our package we would save all of the above into a single file.

### Running Tests

To run the tests in a package, we use the `test()` function from the `devtools` package.

```{r run_tests}
devtools::test()
```

## Behavior-driven development (BDD)

To implement BDD in R, one should use the `testthat::describe()` function. Examples can be seen in `?testthat::describe`.

More robust testing is done using the `testthat` package, however.